code_analyze_unrestrained

### 2025-05-12 code_analyze_unrestrained.qmd
updated code

loc_labels <- c("Non-Sitter", "Sitter", "Crawler", "Walker","Non-Crawler")
loc_colors <- c("tomato","chartreuse4","cyan3","mediumpurple1","orange") %>% set_names(loc_labels)
age_labels <- c("Younger","Older")
age_colors <- c("red","navy") %>% set_names(age_labels)

@@

# outliers
# 178/3, younger min, 0.15
# 148/2, younger max, 0.96
# 181/3, older min, 0.75
outliers_list_older <- 
  DescTools::Outlier(ds_sum$restrained_prop[ds_sum$age_group=="Older"], method = "boxplot")
outliers_list_younger <- 
  DescTools::Outlier(ds_sum$restrained_prop[ds_sum$age_group=="Younger"], method = "boxplot")
ds_sum_excl <- 
  ds_sum[!(ds_sum$restrained_prop %in% outliers_list_older | ds_sum$restrained_prop %in% outliers_list_younger),]

@@

# save the data frames so that don't have to download data from server again
# save(ds, ds_sum, synced_ppts, id_info, id_demo, ds_human, ds_sum_excl, file = "whole-model-prediction.RData")
```

@@

```{r}
ggplot(ds_sum, aes(x = age_group, y = unrestrained_prop)) +
  geom_boxplot() + 
  geom_point(position = position_jitter(width = .1), size = 4, color = "red", shape = 1) + 
  scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))+
  labs(x="Age group",
       y="Unrestrained time %")
ggsave("figure/unrestrained time prop.png", width=6, height=5)

@@

## Q2: Age-related change

```{r}
# across 4~14
ggplot(ds_sum) +
  geom_point(aes(x = agemo, y = unrestrained_prop)) + 
  geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
  scale_x_continuous(name = "Age (mos)", breaks = c(4,7,11,14), limits = c(3,15)) +
  scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))
summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum))

ggplot(ds_sum_excl) +
  geom_point(aes(x = agemo, y = unrestrained_prop)) + 
  geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
  scale_x_continuous(name = "Age (mos)", breaks = c(4,7,11,14), limits = c(3,15)) +
  scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))
ggsave("figure/age_related_change.png",width=6, height=4)
ggsave("figure/age_related_change.eps",width=6, height=4)
summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum_excl))

@@

```{r}
# within younger group comparison
ds_sum %>% 
  filter(age_group=="Younger") %>% 
  ggplot() +
  geom_point(aes(x = agemo, y = unrestrained_prop)) + 
  geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
  scale_x_continuous(name = "Age (mos)", breaks = c(4,7,11,14), limits = c(3,15)) +
  scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))
summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum %>% filter(age_group=="Younger")))

ds_sum_excl %>% 
  filter(age_group=="Younger") %>% 
  ggplot() +
  geom_point(aes(x = agemo, y = unrestrained_prop)) + 
  geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
  scale_x_continuous(name = "Age (mos)", breaks = c(4,7,11,14), limits = c(3,15)) +
  scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))
ggsave("figure/age_related_change_young.png",width=6, height=4)
ggsave("figure/age_related_change_young.eps",width=6, height=4)
summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum_excl%>% filter(age_group=="Younger")))
```

```{r}
# within older group comparison
ds_sum %>% 
  filter(age_group=="Older") %>% 
  ggplot() +
  geom_point(aes(x = agemo, y = unrestrained_prop)) + 
  geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
  scale_x_continuous(name = "Age (mos)", breaks = c(4,7,11,14), limits = c(3,15)) +
  scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))
summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum %>% filter(age_group=="Older")))

ds_sum_excl %>% 
  filter(age_group=="Older") %>% 
  ggplot() +
  geom_point(aes(x = agemo, y = unrestrained_prop)) + 
  geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
  scale_x_continuous(name = "Age (mos)", breaks = c(4,7,11,14), limits = c(3,15)) +
  scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))
ggsave("figure/age_related_change_old.png",width=6, height=4)
ggsave("figure/age_related_change_old.eps",width=6, height=4)
summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum_excl%>% filter(age_group=="Older")))
```

@@

## Q3-1: sitting related change
```{r}
ds_sum %>% 
  ggplot() +
    geom_point(aes(x = agemo, y = unrestrained_prop, color = sitter)) + 
    geom_smooth(aes(x = agemo, y = unrestrained_prop, color = sitter), method = "lm", se = FALSE) + 
    scale_x_continuous(name = "Age (mo)", breaks = c(4,7,11,14), limits = c(3,15)) +
    scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))+
    scale_color_manual(values=loc_colors)
ggsave("figure/sit_related_change.png", width=6, height=4)
ggsave("figure/sit_related_change.eps", width=6, height=4)
summary(lm(unrestrained_prop ~ sitter + scale(agemo), data = ds_sum))
summary(lm(unrestrained_prop ~ sitter + scale(agemo), data = ds_sum_excl))

# ds_sum %>% 
#   filter(age_group=="Younger") %>% 
#   ggplot() +
#     geom_point(aes(x = agemo, y = unrestrained_prop, color = sitter)) + 
#     geom_smooth(aes(x = agemo, y = unrestrained_prop, color = sitter), method = "lm", se = FALSE) + 
#     scale_x_continuous(name = "Age (mo)", breaks = c(4,7,11,14), limits = c(3,15)) +
#     scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))+
#     scale_color_manual(values=loc_colors)
summary(lm(unrestrained_prop ~ sitter + scale(agemo), data = ds_sum %>% filter(age_group=="Younger")))
# summary(lm(unrestrained_prop ~ sitter + scale(age_base)+scale(age_change), data = ds_sum2 %>% filter(age_group=="Younger")))

@@

## Q3-2: crawling related change
```{r}
ds_sum %>% 
  ggplot() +
    geom_point(aes(x = agemo, y = unrestrained_prop, color = crawler)) + 
    geom_smooth(aes(x = agemo, y = unrestrained_prop, color = crawler), method = "lm", se = FALSE) + 
    scale_x_continuous(name = "Age (mo)", breaks = c(4,7,11,14), limits = c(3,15)) +
    scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))+
    scale_color_manual(values=loc_colors)
ggsave("figure/crawl_related_change.png", width=6, height=4)
ggsave("figure/crawl_related_change.eps", width=6, height=4)
summary(lm(unrestrained_prop ~ crawler + scale(agemo), data = ds_sum))
summary(lm(unrestrained_prop ~ crawler + scale(agemo), data = ds_sum_excl))

## Q3-3: crawling + sitting
```{r}
ds_sum %>% 
  mutate(loc = ifelse(loc=="Walker","Crawler",loc)) %>% 
  ggplot() +
    geom_point(aes(x = agemo, y = unrestrained_prop, color = loc)) + 
    geom_smooth(aes(x = agemo, y = unrestrained_prop, color = loc), method = "lm", se = FALSE) + 
    scale_x_continuous(name = "Age (mo)", breaks = c(4,7,11,14), limits = c(3,15)) +
    scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))+
    scale_color_manual(values=loc_colors)
ggsave("figure/crawl_sit_change.png", width=6, height=4)
ggsave("figure/crawl_sit_change.eps", width=6, height=4)
summary(lm(unrestrained_prop ~ crawler + sitter + agemo, data = ds_sum))
summary(lm(unrestrained_prop ~ crawler + sitter + agemo, data = ds_sum_excl))
# here, excluding outliers or not influences a lot
```

>>>>

loc_labels <- c("Non-Sitter", "Sitter", "Crawler", "Walker","Non-Crawler")
loc_colors <- c("#D81B60","#004D40","#1E88E5","purple4","orange") %>% set_names(loc_labels)
age_labels <- c("Younger","Older")
age_colors <- c("red","navy") %>% set_names(age_labels)

@@

# outliers
# 178/3, younger min, 0.15
# 148/2, younger max, 0.96
# 181/3, older min, 0.75
outliers_list_older <- 
  DescTools::Outlier(ds_sum$unrestrained_prop[ds_sum$age_group=="Older"], method = "boxplot")
outliers_list_younger <- 
  DescTools::Outlier(ds_sum$unrestrained_prop[ds_sum$age_group=="Younger"], method = "boxplot")
ds_sum_excl <- 
  ds_sum[!(ds_sum$unrestrained_prop %in% outliers_list_older | ds_sum$unrestrained_prop %in% outliers_list_younger),]

outliers_list <- 
  DescTools::Outlier(ds_sum$unrestrained_prop, method = "boxplot")
ds_sum_excl2 <- 
  ds_sum[!(ds_sum$unrestrained_prop %in% outliers_list),]

@@

# save the data frames so that don't have to download data from server again
# save(ds, ds_sum, synced_ppts, id_info, id_demo, ds_human, ds_sum_excl, ds_sum_excl2, file = "whole-model-prediction.RData")
```

@@

```{r}
ggplot(ds_sum, aes(x = age_group, y = unrestrained_prop)) +
  geom_boxplot() + 
  geom_point(position = position_jitter(width = .1), size = 4, color = "red", shape = 1) + 
  scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))+
  labs(x="Age group",
       y="Unrestrained time %")
ggsave("figure/unrestrained time prop.png", width=6, height=5)

@@

## Q2: Age-related change
```{r}
# descriptive results
ds_sum %>% 
  group_by(age_mo) %>% 
  summarize(mean = mean(unrestrained_prop),
            sd = sd(unrestrained_prop),
            .groups="drop")
```

```{r}
# across 4~14
# ggplot(ds_sum) +
#   geom_point(aes(x = agemo, y = unrestrained_prop)) + 
#   geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
#   scale_x_continuous(name = "Age (mo.)", breaks = c(4,7,11,14), limits = c(3,15)) +
#   scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))
summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum))

ggplot(ds_sum) +
  geom_point(aes(x = agemo, y = unrestrained_prop)) + 
  geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
  scale_x_continuous(name = "Age (mo.)", breaks = c(4,7,11,14), limits = c(3,15)) +
  scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))
ggsave("figure/age_related_change.png",width=6, height=4)
ggsave("figure/age_related_change.eps",width=6, height=4)

ggplot(ds_sum_excl2) +
  geom_point(aes(x = agemo, y = unrestrained_prop)) + 
  geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
  scale_x_continuous(name = "Age (mo.)", breaks = c(4,7,11,14), limits = c(3,15)) +
  scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))
ggsave("figure/age_related_change.png",width=6, height=4)
ggsave("figure/age_related_change.eps",width=6, height=4)
summary(lm(unrestrained_prop ~ agemo, data = ds_sum_excl2))

ggplot(ds_sum_excl2) +
  geom_point(aes(x = agemo, y = unrestrained_prop),size=4, shape=1, stroke=1.5) +
  geom_smooth(aes(x = agemo, y = unrestrained_prop, group=age_group), method = "lm", se = T) +
  scale_x_continuous(name = "Age (mo.)", breaks = c(4,7,11,14), limits = c(3,15)) +
  scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))
ggsave("figure/age_related_change_ingroup.png",width=6, height=4)
ggsave("figure/age_related_change_ingroup.eps",width=6, height=4)

@@

```{r}
# within younger group comparison
ds_sum %>% 
  filter(age_group=="Younger") %>% 
  ggplot() +
  geom_point(aes(x = agemo, y = unrestrained_prop)) + 
  geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
  scale_x_continuous(name = "Age (mo.)", breaks = c(4,7,11,14), limits = c(3,15)) +
  scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))
summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum %>% filter(age_group=="Younger")))

ds_sum_excl2 %>%
  filter(age_group=="Younger") %>%
  ggplot() +
  geom_point(aes(x = agemo, y = unrestrained_prop)) +
  geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) +
  scale_x_continuous(name = "Age (mo.)", breaks = c(4,7,11,14), limits = c(3,15)) +
  scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))
ggsave("figure/age_related_change_young.png",width=6, height=4)
ggsave("figure/age_related_change_young.eps",width=6, height=4)

summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum_excl2%>% filter(age_group=="Younger")))
```

```{r}
# within older group comparison
ds_sum %>% 
  filter(age_group=="Older") %>% 
  ggplot() +
  geom_point(aes(x = agemo, y = unrestrained_prop)) + 
  geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
  scale_x_continuous(name = "Age (mo.)", breaks = c(4,7,11,14), limits = c(3,15)) +
  scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))
summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum %>% filter(age_group=="Older")))

ds_sum_excl2 %>% 
  filter(age_group=="Older") %>% 
  ggplot() +
  geom_point(aes(x = agemo, y = unrestrained_prop)) + 
  geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
  scale_x_continuous(name = "Age (mo.)", breaks = c(4,7,11,14), limits = c(3,15)) +
  scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))
ggsave("figure/age_related_change_old.png",width=6, height=4)
ggsave("figure/age_related_change_old.eps",width=6, height=4)
summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum_excl2%>% filter(age_group=="Older")))
```

@@

## Q3 locomotor-related change: crawling + sitting( + walking)
```{r}
ds_sum %>% 
  mutate(loc = ifelse(loc=="Walker","Crawler",loc))%>%
  mutate(loc = factor(loc, 
                      levels = c("Non-Sitter","Sitter","Crawler"), 
                      labels=c("Non-Sitter","Sitter","Crawler")))%>% 
  ggplot() +
    geom_point(aes(x = agemo, y = unrestrained_prop, color = loc)) + 
    scale_x_continuous(name = "Age (mo)", breaks = c(4,7,11,14), limits = c(3,15)) +
    scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))+
    scale_color_manual(values=loc_colors)

# sitting + crawling
ds_sum_excl2 %>% 
  mutate(loc = ifelse(loc=="Walker","Crawler",loc))%>%
  mutate(loc = factor(loc, 
                      levels = c("Non-Sitter","Sitter","Crawler"), 
                      labels=c("Non-Sitter","Sitter","Crawler")))%>% 
  ggplot() +
    geom_point(aes(x = agemo, y = unrestrained_prop, color = loc)) + 
    scale_x_continuous(name = "Age (mo)", breaks = c(4,7,11,14), limits = c(3,15)) +
    scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))+
    scale_color_manual(values=loc_colors)

# sitting + crawling + walking
mean_unr_loc <- ds_sum_excl2 %>% 
  group_by(loc) %>% 
  summarize(age = mean(agemo),
            prop = mean(unrestrained_prop),
            .groups="drop") %>% 
  mutate(x=age-0.8,
         xend=age+0.8,
         y=prop,
         yend=prop,)

ds_sum_excl2 %>% 
  ggplot() +
    geom_point(aes(x = agemo, y = unrestrained_prop, color = loc),size=4,shape=1, stroke=1.5) + 
    scale_x_continuous(name = "Age (mon)", breaks = c(4,7,11,14), limits = c(3,15)) +
    scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))+
    scale_color_manual(values=loc_colors)
ggsave("figure/sit_crawl_walk_change.png", width=8, height=4)
ggsave("figure/sit_crawl_walk_change.eps", width=8, height=4)
# summary(lm(unrestrained_prop ~ crawler + sitter + agemo, data = ds_sum))
# summary(lm(unrestrained_prop ~ crawler + sitter + agemo, data = ds_sum_excl))

summary(lm(unrestrained_prop ~ agemo, data = ds_sum_excl2))
summary(lm(unrestrained_prop ~ sitter + agemo, data = ds_sum_excl2))
summary(lm(unrestrained_prop ~ sitter + crawler + agemo, data = ds_sum_excl2))
summary(lm(unrestrained_prop ~ sitter + crawler + walker + agemo, data = ds_sum_excl2))
```

```{r}
# description
t <- ds_sum_excl2 %>% 
  group_by(loc) %>% 
  summarize(mean=mean(unrestrained_prop),
            sd=sd(unrestrained_prop))

@@
